\documentclass[a4paper, 12pt]{article}

\usepackage{amsmath, amsfonts, enumitem, fullpage, hyperref}

\newcommand{\realR}{\mathbb{R}}
\newcommand{\ratQ}{\mathbb{Q}}
\newcommand{\intZ}{\mathbb{Z}}
\newcommand{\Holder}{H\"{o}lder}

\title{MA3209, Homework 1}

\date{}
\begin{document}

\maketitle

Due at the beginning of your tutorial session, Monday 27 August or Wednesday 29 August.

\begin{enumerate}
\item
  Consider the metric spaces $(\realR^n, \rho_p(x, y) = (\sum^n_{k=1} \lvert x_k - y_k \rvert^p)^{1/p})$ and $(\realR^n, \rho_{\infty}(x, y) = \max^n_{k=1} \lvert x_k - y_k \rvert)$.
  \begin{enumerate}
  \item
    Show that the function $\rho_{\infty}(x, y)$ satisfies the triangle axiom.
  \item
    Show that for any $p \geq 1$ and $x, y$ in $\realR^n$
    \begin{equation*}
      \rho_{\infty}(x, y) \leq \rho_p(x, y) \leq n \rho_{\infty}(x, y).
    \end{equation*}
  \item
    Show that for any $p > 1$ and $x, y$ in $\realR^n$
    \begin{equation*}
      \rho_p(x, y) \leq \rho_1(x, y) \leq n^{1 - \frac{1}{p}} \rho_p(x, y).
    \end{equation*}
    (Hint: For the left inequality, show $a^p + b^p \leq (a + b)^p$ for all $a, b > 0$ first. For the right inequality, use \Holder's inequality.)
  \end{enumerate}
\item
  Show that the distance function in metric space $C^p([c_1, c_2])$ 
  \begin{equation*}
    \rho_p(x(t), y(t)) = \left( \int^{c_2}_{c_1} \lvert x(t) - y(t) \rvert^p \right)^{\frac{1}{p}}
  \end{equation*}
  satisfies the triangle inequality when $p > 1$, in the following steps.
  \begin{enumerate}
  \item
    Show that the triangle inequality follows from the integral form of the Minkowski inequality for any continuous functions $a(t), b(t)$ on $[c_1, c_2]$
    \begin{equation} \label{eq:int_Minkowski}
      \left( \int^{c_2}_{c_1} \lvert a(t) + b(t) \rvert^p dt \right)^{\frac{1}{p}} \leq \left( \int^{c_2}_{c_1} \lvert a(t) \rvert^p dt \right)^{\frac{1}{p}} + \left( \int^{c_2}_{c_1} \lvert b(t) \rvert^p dt \right)^{\frac{1}{p}}.
    \end{equation}
  \item
    Show that the integral form of the Minkowski inequality \eqref{eq:int_Minkowski} follows from the integral form of \Holder's inequality for any continuous functions $x(t), y(t)$ on $[c_1, c_2]$, and $p, q > 1$ such that $\frac{1}{p} + \frac{1}{q} = 1$
    \begin{equation} \label{eq:int_Holder}
      \int^{c_2}_{c_1} \lvert x(t)y(t) \rvert dt \leq \left( \int^{c_2}_{c_1} \lvert x(t) \rvert^p dt \right)^{\frac{1}{p}} \left( \int^{c_2}_{c_1} \lvert y(t) \rvert^q dt \right)^{\frac{1}{q}}.
    \end{equation}
    (Hint: Use the trick
    \begin{equation*}
      \int^{c_2}_{c_1} \lvert a(t) + b(t) \rvert^p dt = \int^{c_2}_{c_1} \lvert a(t) + b(t) \rvert^{p-1} \lvert a(t) \rvert dt + \int^{c_2}_{c_1} \lvert a(t) + b(t) \rvert^{p-1} \lvert b(t) \rvert dt.
    \end{equation*}
)
\item
  Prove the integral form of \Holder's inequality \eqref{eq:int_Holder}.

  (Hint: Reduce it to the special case that $\int^{c_2}_{c_1} \lvert x(t) \rvert^p dt = 1$, $\int^{c_2}_{c_1} \lvert y(t) \rvert^q dt = 1$.)
  \end{enumerate}
\item
  Let $(X, \rho)$ be a metric space. Define the function $d: X \times X \to \realR$ by
  \begin{equation*}
    d(x, y) = \frac{\rho(x, y)}{1 + \rho(x, y)}.
  \end{equation*}
  Show that $(X, d)$ is also a metric space.
\item
  Let $V$ be a vector space, and let $\lVert \cdot \rVert: V \to \realR$ be a function (called a \emph{norm} on $V$) satisfying the following conditoins:
  \begin{enumerate}[label=(\roman{*})]
  \item
    $\lVert x \rVert \geq 0$ for all $x$ in $V$.
  \item
    $\lVert x \rVert = 0$ if and only if $x = 0$.
  \item
    $\alpha \lVert x \rVert = \lvert \alpha \rvert \lVert x \rVert$ for all $x$ in $V$ and all scalar $\alpha$.
  \item
    $\lVert x + y \rVert \leq \lVert x \rVert + \lVert y \rVert$ for all $x, y$ in $V$.
  \end{enumerate}
  Define $\rho: V \times V \to \realR$ by $\rho(x, y) = \lVert x - y \rVert$ for $x, y$ in $V$. Show that $(V, \rho)$ is a metric space.
\item (Extra credit)
  Let $p$ be a prime number. For any integer $n > 0$ we define $v_p(n)$ as the exponent of $p$ in the decomposition of $n$ into prime numbers. Let $x = \pm \frac{r}{s}$ be any rational number $\neq 0$, with $r$ and $s$ integers $> 0$. Define
  \begin{equation} \label{eq:defn_of_v_p_on_rationals}
    v_p(x) = v_p(r) - v_p(s).
  \end{equation}
  Then define the function $\rho: \ratQ \times \ratQ \to \realR$ by ($\ratQ$ stands for the set of rational numbers)
  \begin{equation*}
    \rho(x, y) =
    \begin{cases}
      p^{-v_p(x - y)} & \text{if $x \neq y$,} \\
      0 & \text{if $x = y$.}
    \end{cases}
  \end{equation*}
    \begin{enumerate}
    \item
      Show that $v_p(x)$ is well defined, that is, $v_p(x)$ in \eqref{eq:defn_of_v_p_on_rationals} does not depend on the particular expression of $x$ as a fraction.
    \item
      Show that $(\ratQ, \rho)$ is a metric space. ($\rho$ is the $p$-adic distance useful in number theory.)
    \end{enumerate}
\end{enumerate}

\end{document}
