% Jason R. Blevins - Curriculum Vitae
%
% Copyright (C) 2004-2011 Jason R. Blevins
%
% You may use use this document as a template to create your own CV
% and you may redistribute the source code freely.  No attribution is
% required in any resulting documents.  I do ask that you please leave
% this notice and the above URL in the source code if you choose to
% redistribute this file.

\documentclass[12pt,a4paper]{article}

\usepackage{url, hyperref}
\usepackage{geometry}
\usepackage{etaremune}
\usepackage{amsmath}

% Fonts
%\usepackage[T1]{fontenc}
%\usepackage[urw-garamond]{mathdesign}
\usepackage{mathptmx}

\hyphenation{Chang-chun}

% Set your name here
\def\name{Dong Wang}

% The following metadata will show up in the PDF properties
\hypersetup{
  colorlinks = true,
  urlcolor = black,
  pdfauthor = {\name},
  pdfkeywords = {random matrix theory, integrable system, probability},
  pdftitle = {\name: Curriculum Vitae},
  pdfsubject = {Curriculum Vitae},
  pdfpagemode = UseNone
}

\geometry{
  body={6.5in, 9.0in},
  left=1.0in,
  top=1.0in
}

% Customize page headers
\pagestyle{myheadings}
\markright{\name}
\thispagestyle{empty}

% Custom section fonts
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\Large}
\subsectionfont{\rmfamily\mdseries\itshape\large}

% Other possible font commands include:
% \ttfamily for teletype,
% \sffamily for sans serif,
% \bfseries for bold,
% \scshape for small caps,
% \normalsize, \large, \Large, \LARGE sizes.

% Don't indent paragraphs.
\setlength\parindent{0em}

%Make lists without bullets and compact spacing
\newenvironment{item_list}{
 \begin{list}{}{
   \setlength{\leftmargin}{1.5em}
   \setlength{\itemsep}{0.25em}
   \setlength{\parskip}{0pt}
   \setlength{\parsep}{0.25em}
 }
}{
 \end{list}
}

\begin{document}

% Place name at left
{\huge \name}

% Alternatively, print name centered and bold:
%\centerline{\huge \bf \name}

\bigskip

\begin{minipage}[t]{0.5\textwidth}
  Department of Mathematics \\
  National University of Singapore \\
  Block S17 6-20 \\
  10 Lower Kent Ridge Road \\
  Singapore, 119076
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
  Phone: +65 6516 2746 \\
  Fax: +65 6779 5452 \\
  %Office: Block S17 6-20 \\
  Email: \href{mailto:matwd@nus.edu.sg}{matwd@nus.edu.sg} \\
  Homepage: \href{https://wangd-math.github.io/}{{https://wangd-math.github.io/}}
\end{minipage}

\section*{Education}

\begin{item_list}
  \item Ph.D. Mathematics, Brandeis University, 2008.
    % Received May 16, 2010
    \begin{item_list}
    \item \textit{Advisor:}
      Mark Adler.
    \end{item_list}
  \item B.S. Mathematics, %\textit{Summa Cum Laude},
    Peking University, 2003.
    % Received May 14, 2004
  %\item Budapest Semesters in Mathematics, Fall 2002.
\end{item_list}

\section*{Work Experience}

\begin{item_list}
\item Assistant Professor, Department of Mathematics, National University of Singapore, Auguest 2012--present.
\item Post-Doc Assistant Professor, Department of Mathematics, University of Michigan,
  July 2009--June 2012.
\item Postdoctoral Researcher, Institut de recherche en math\'{e}matique et physique (IRMP), Universit\'{e} catholique de Louvain, 2011, March--August.
\item Postdoctoral Fellow, Centre de Recherches Math\'{e}matiques (CRM), Universit\'{e} de Montr\'{e}al, 2008--2009.
\item Teaching Assistant, Department of Mathematics, Brandeis University, 2003--2008.
\end{item_list}

\section*{Research}

I am an analyst working on probability problems, and the proper term describing my research interest is ``integrable probability''. My research focuses on random matrix theory and interacting particle systems. I am also interested in integrable systems, multiple orthogonal polynomials and Riemann-Hilbert problems related to random matrices.

\section*{In Preparation}

\begin{itemize}
\item
  Dong Wang, The largest eigenvalue of real symmetric, Hermitian and Hermitian self-dual random matrix models with rank one external source, part II.
\end{itemize}

\section*{Papers and Preprints}

\begin{etaremune}
\item
  Swapnil Yadav, Kazi Alam, Khandker Muttalib and Dong Wang, \href{https://arxiv.org/abs/1908.03726}{Generalized random matrix model with additional interactions}, \href{https://dx.doi.org/10.1088/1751-8121/ab56e0}{\textit{J.\ Phys.\ A}, 53 (2020), no.\ 1, 015001, 15 pp}.
\item
  Zhipeng Liu, Axel Saenz and Dong Wang, \href{https://arxiv.org/abs/1905.02987}{Integral formulas of ASEP and $q$-TAZRP on a ring}, \url{https://arxiv.org/abs/1905.02987}.
\item
  Dang-Zheng Liu, Dong Wang and Yanhui Wang, \href{https://arxiv.org/abs/1810.00433}{Lyapunov exponent, universality and phase transition for products of random matrices}, \url{https://arxiv.org/abs/1810.00433}.
\item
  Karl Liechty and Dong Wang, \href{https://arxiv.org/abs/1706.06653}{Asymptotics of free fermions in a quadratic well at finite temperature and the Moshe-Neuberger-Shapiro random matrix model}, \href{https://doi.org/10.1214/19-AIHP994}{\textit{Ann.\ Inst.\ Henri Poincar\'{e} Probab.\ Stat.}, 56 (2020), no.\ 2, 1072--1098}.
\item
  Eunghyun Lee and Dong Wang, \href{https://arxiv.org/abs/1703.08839}{Distributions of a particle's position and their asymptotics in the $q$-deformed totally asymmetric zero range process with site dependent jumping rates}, \href{https://doi.org/10.1016/j.spa.2018.06.005}{\textit{Stochastic Process.\ Appl.}, 129 (2019), no.\ 5, 1795--1828}.
\item
  Karl Liechty and Dong Wang, \href{http://arxiv.org/abs/1608.08712}{Nonintersecting Brownian bridges between reflecting or absorbing walls}, \href{https://doi.org/10.1016/j.aim.2016.10.024}{\textit{Adv.\ Math.}, 309 (2017), 155--208}.
\item
  Tom Claeys, Arno B.~J.~Kuijlaars, Karl Liechty and Dong Wang, \href{http://arxiv.org/abs/1608.05870}{Propagation of singular behavior for Gaussian perturbations of random matrices}, \href{https://doi.org/10.1007/s00220-018-3195-8}{\textit{Comm.\ Math.\ Phys.}, 362 (2018), no.\ 1, 1--54}.
\item
  Karl Liechty and Dong Wang, \href{http://arxiv.org/abs/1601.01603}{Two Lax systems for the Painlev\'{e} II equation, and two related kernels in random matrix theory}, \href{https://doi.org/10.1137/16M1056080}{\textit{SIAM J.\ Math.\ Anal.}, 48 (2016), no.\ 5, 3618--3666}.
\item
  Dong Wang and David Waugh, \href{http://arxiv.org/abs/1512.01612}{The transition probability of the $q$-TAZRP ($q$-Bosons) with inhomogeneous jump rates}, \href{https://doi.org/10.3842/SIGMA.2016.037}{\textit{SIGMA} 12 (2016), 036, 16 pp, Contribution to the special issue on Asymptotics and Universality in Random Matrices, Random Growth Processes, Integrable Systems and Statistics in honor of Percy Deift and Craig Tracy.}
\item 
  Tom Claeys, Arno B.~J.~Kuijlaars and Dong Wang, \href{http://arxiv.org/abs/1505.00610}{Correlation kernels for sums and products of random matrices}, \href{https://doi.org/10.1142/S2010326315500173}{\textit{Random Matrices Theory Appl.}, 4 (2015), no.\ 4, 1550017, 31pp}.
\item
  Peter J.~Forrester and Dong Wang, \href{http://arxiv.org/abs/1502.07147}{Muttalib--Borodin ensembles in random matrix theory --- realisations and correlation functions}, \href{https://doi.org/10.1214/17-EJP62}{\textit{Electron. J. Probab.}, 22 (2017), paper no.\ 54, 43pp}.
\item
  Dang-Zheng Liu, Dong Wang and Lun Zhang, \href{http://arxiv.org/abs/1412.6777}{Bulk and soft-edge universality for singular values of products of Ginibre random matrices}, \href{https://doi.org/10.1214/15-AIHP696}{\textit{Ann. Inst. Henri Poincar\'e Probab. Stat.}, 52 (2016), no.\ 4, 1734--1762}.
\item
  Ivan Corwin, Zhipeng Liu and Dong Wang, \href{http://arxiv.org/abs/1412.5087}{Fluctuations of TASEP and LPP with general initial data}, \href{https://doi.org/10.1214/15-AAP1139}{\textit{Ann.\ Appl.\ Probab.}, 26 (2016), no.\ 4, 2030--2082}.
\item
  Karl Liechty and Dong Wang, \href{http://arxiv.org/abs/1312.7390}{Nonintersecting Brownian motions on the unit circle}, \href{https://doi.org/10.1214/14-AOP998}{\textit{Ann.\ Probab.}, 44 (2016), no.\ 2, 1134--1211}.
\item 
  Mark Adler, Pierre van Moerbeke and Dong Wang,
  \href{http://arxiv.org/abs/1301.7017}{Random matrix minor processes related to percolation theory}, \href{https://doi.org/10.1142/S2010326313500081}{\textit{Random Matrices Theory Appl.}, 2 (2014), no.\ 4, 135008, 72pp}.
\item 
  Tom Claeys and Dong Wang,
  \href{http://arxiv.org/abs/1212.3768}{Random matrices with equispaced external source}, \href{https://doi.org/10.1007/s00220-014-1988-y}{\textit{Comm.\ Math.\ Phys.}, 328 (2014), no.\ 3, 1023--1077}.
\item Jinho Baik and Dong Wang,
  \href{http://arxiv.org/abs/1207.0389}{On a relationship between high rank cases and rank one cases of Hermitian random matrix models with external source}, \href{http://www.cambridge.org/asia/catalogue/catalogue.asp?isbn=9781107079922}{\textit{Random Matrix Theory, Interacting Particle Systems and Integrable Systems}, Edited by Percey Deift and Peter Forrester, MSRI Publications 65 (2014), Cambridge University Press, Cambridge, 25--38}.
\item Jinho Baik and Dong Wang,
  \href{http://arxiv.org/abs/1104.2915}{On the largest eigenvalue of a Hermitian random matrix model with spiked external source II. Higher rank case}, \href{https://doi.org/10.1093/imrn/rns136}{\textit{Int.\ Math.\ Res.\ Not.\ IMRN}, (2013) no.\ 14, 3304--3370}.
\item Dong Wang,
  \href{http://arxiv.org/abs/1012.4144}{The largest eigenvalue of real symmetric, Hermitian and Hermitian self-dual random matrix models with rank one external source, part I}, \href{https://doi.org/10.1007/s10955-012-0417-x}{\textit{J. Stat.\ Phys.}, 146 (2012) no.\ 4, 719--761}.
\item Jinho Baik and  Dong Wang,
  \href{http://arxiv.org/abs/1010.4604}{On the largest eigenvalue of a Hermitian random matrix model with spiked external source I. Rank 1 case}, \href{https://doi.org/10.1093/imrn/rnq257}{\textit{Int.\ Math.\ Res.\ Not.\ IMRN}, (2011) no.\ 22, 5164--5240}.
\item Dong Wang,
  \href{http://arxiv.org/abs/0810.0280}{Random matrices with external source and KP $\tau$ functions}, \href{https://doi.org/10.1063/1.3167802}{\textit{J.\ Math.\ Phys.}, 50 (2009), no.\ 7, 073506, 10pp}.
\item Dong Wang,
 \href{http://arxiv.org/abs/0711.2722}{The largest sample eigenvalue distribution in the rank 1 quaternionic spiked model of Wishart ensemble}, \href{https://doi.org/10.1214/08-AOP432}{\textit{Ann.\ Probab.}, 37 (2009), no.\ 4, 1273--1328}.
\item Dong Wang,
  \href{http://arxiv.org/abs/0711.3797}{A PDE for the multi-time joint probability of the Airy process}, \href{https://doi.org/10.1016/j.physd.2009.02.007}{\textit{Phys.\ D}, 238 (2009), no.\ 8, 819--833}.
\end{etaremune}

\subsection*{Dissertation}
\begin{itemize}
\item
  Spiked Models in Wishart Ensemble.
\end{itemize}

\section*{Teaching}

\subsection*{{National University of Singapore}}

\begin{item_list}
\item
  Complex Analysis I, \href{https://wangd-math.github.io/teaching/MA3111_S20/index.html}{MA3111}, 2019/2020 Semester 2, (and two tutorials).
\item
  Complex Analysis I, \href{https://wangd-math.github.io/teaching/MA3111_F19/index.html}{MA3111}, 2019/2020 Semester 1, (and two tutorials).
\item
  Complex Analysis I, \href{https://wangd-math.github.io/teaching/MA3111_S19/index.html}{MA3111}, 2018/2019 Semester 2, (and three tutorials).
\item
  Complex Analysis I, \href{https://wangd-math.github.io/teaching/MA3111_F18/index.html}{MA3111}, 2018/2019 Semester 1, (and two tutorials).
\item
  Functional Analysis, \href{https://wangd-math.github.io/teaching/MA4211_S18/index.html}{MA4211}, 2017/2018 Semester 2, (and a tutorial).
\item
  Measure and Integration, \href{https://wangd-math.github.io/teaching/MA4262_F17/index.html}{MA4262}, 2017/2018, Semester 1, (and a tutorial).
\item
  Stochastic Processes I, \href{https://wangd-math.github.io/teaching/MA3238_ST3236_S17/index.html}{MA3238/ST3236}, 2016/2017, Semester 2, (and four tutorials).
\item
  Measure and Integration, \href{https://wangd-math.github.io/teaching/MA4262_F16/index.html}{MA4262}, 2016/2017, Semester 1, (and a tutorial).
\item
  Stochastic Processes I, \href{https://wangd-math.github.io/teaching/MA3238_ST3236_S16/index.html}{MA3238/ST3236}, 2015/2016, Semester 2, (and five tutorials).
\item
  Probability Theory I, \href{https://wangd-math.github.io/teaching/MA5259_F15/index.html}{MA5259}, 2015/2016 Semester 1.
\item
  Stochastic Processes I, \href{https://wangd-math.github.io/teaching/MA3238_ST3236_S15/index.html}{MA3238/ST3236}, 2014/2015 Semester 2, (and six tutorials).
\item
  Mathematical Analysis III, \href{https://wangd-math.github.io/teaching/MA3209_F14/index.html}{MA3209}, 2014/2015 Semester 1, (and two tutorials).
\item
  Functional Analysis, \href{https://wangd-math.github.io/teaching/MA4211_S14/index.html}{MA4211}, 2013/2014 Semester 2, (and a tutorial).
\item
  Mathematical Analysis III, \href{https://wangd-math.github.io/teaching/MA3209_F13/index.html}{MA3209}, 2013/2014 Semester 1, (and two tutorials).
\item
  Topics in Random Matrix Theory, \href{https://wangd-math.github.io/teaching/MA6252_S13/index.html}{MA6252}, 2012/2013 Semester 2.
\item
  Mathematical Analysis III, \href{https://wangd-math.github.io/teaching/MA3209_F12/index.html}{MA3209}, 2012/2013 Semester 1, (and two tutorials).
\end{item_list}

\subsection*{University of Michigan}

\begin{item_list}
\item Multivariable Calculus, \href{http://instruct.math.lsa.umich.edu/}{Math 217}, Spring 2012,
\item Matrix Algebra, \href{http://www.math.lsa.umich.edu/courses/undergrad.html#417}{Math 417}, Winter 2012 (two sections),
\item Integral Caculus, \href{http://www.math.lsa.umich.edu/courses/116/}{Math 116}, Fall 2011 (two sections),
\item Differential Calculus, \href{http://www.math.lsa.umich.edu/courses/115/}{Math 115}, Fall 2010 (two sections). 
\item Differential Calculus, \href{http://www.math.lsa.umich.edu/courses/115/}{Math 115}, Winter 2010 (two sections). 
\item Differential Calculus, \href{http://www.math.lsa.umich.edu/courses/115/}{Math 115}, Fall 2009 (two sections).
\end{item_list}

\subsection*{Brandeis University}

\begin{item_list}
\item Differential Calculus, Math 10a,
  Fall 2007. 
\item Differential Calculus, Math 10a,
  Spring 2007. 
\item Integral Calculus, Math 10b,
  Fall 2006.
\item Differential Calculus, Math 10a,
  Spring 2006.
\end{item_list}

\section*{Students mentored}

\subsection*{Master students}

David Waugh (NUS 2015)

\subsection*{Selected undergraduate students}

Mark Ng Say-Yao (NUS Undergraduate Research Opportunities Programme in Science (UROPS), winner of 2015/16 CRISP award)

\section*{Conference and Seminar Presentations}

\begin{item_list}
\item
  \href{https://www.intindstat.org/iisaconference2019/}{INDSTATS: Innovations in Data and Statistical Sciences}, Indian Institute of Technology Bombay, Mumbai, India, December 2019.
\item
  \href{http://www.ims-china.org/}{IMS-China conference}, Dalian University of Technology, Dalian, China, July 2019.
\item
  \href{http://ww1.math.nus.edu.sg/seminars.aspx?CatID=5}{Topology and geometry seminar}, National University of Singapore, Singapore, April 2019.
\item
  \href{https://sites.google.com/view/randommatrices2019nus}{Workshop on random matrices, stochastic geometry and related topics}, Institute for Mathematical Sciences, Singapore, March 2019.
\item
  \href{http://math.fudan.edu.cn/show.aspx?info_lb=766&flag=527&info_id=6117}{Workshop on random matrices and related fields}, Fudan University, Shanghai, China, January 2019.
\item
  \href{https://www.uni-bielefeld.de/(en)/ZiF/AG/2018/08-27-Akemann.html}{Sums and products of random matrices}, Universit\"{a}t Bielefeld, Bielefeld, Germany, August 2018.
\item
  \href{https://ims-aprm2018.stat.nus.edu.sg/}{The $5^{\text{th}}$ Institute of Mathematical Statistics Asia Pacific Rim Meeting (IMS-APRM)}, National University of Singapore, Singapore, June 2018.
\item
  \href{http://math.fudan.edu.cn/en/Data/View/-99994362}{Workshop on integrable systems and orthogonal polynomials}, Fudan University, Shanghai, China, January 2018.
\item
  Working group in integrable systems, Tsinghua University, Beijing, China, December 2017.
\item
  \href{http://wulab.ustc.edu.cn/bencandy.php?fid=61&id=1688}{Workshop on Stochastic Analysis and Random Matrices}, University of Science and Technology of China, Hefei, China, December 2017.
\item
  \href{https://lsa.umich.edu/math/centers-outreach/mcaim/painleve-equations-workshop--august-25-29--2017.html}{Painlev\'{e} Equations and Applications: A Workshop in Memory of A.~A.~Kapaev}, University of Michigan, Ann Arbor, USA, August 2017.
\item
  \href{http://www.fst.umac.mo/en/staff/documents/yayangchen/Workshop%20on%20RMT17.pdf}{Random Matrices, EurAsia 2017}, Fudan University, Shanghai, China, July 2017.
\item
  \href{http://hkumath.hku.hk/~imr/records1617/Workshop%20on%20Random%20Matrices_2017Jun16.pdf}{Workshop on Random Matrices}, Hong Kong University, Hong Kong, China, June 2017.
\item 
  \href{http://www6.cityu.edu.hk/rcms/icsf2017/index.htm}{International Conference on Special Functions: Theory, Computation, and Applications}, City University of Hong Kong, Hong Kong, China, June 2017.
\item
  \href{http://www.math.pku.edu.cn/teachers/liuyong/seminar/wangdong2016.pdf}{Probability Seminar}, Peking University, Beijing, China, December 2016.
\item
  \href{http://www.math.sjtu.edu.cn/conference/2016icsa/}{The $10^{\text{th}}$ ICSA International Conference on Global Growth of Modern Statistics in the $21^{\text{st}}$ Century}, Shanghai Jiao Tong University, Shanghai, China, December 2016.
\item
  \href{https://www2.physik.uni-bielefeld.de/rpm_2016.html}{Random Product Matrices --- New Developments and Applications}, Universit\"{a}t Bielefeld, Bielefeld, Germany, August 2016.
\item
  \href{http://home.ustc.edu.cn/~shanamy3/index.html}{Probability Forum}, University of Science and Technology of China, Hefei, China, August 2016.
\item
  \href{http://www.fst.umac.mo/conference/RandomMatricesEurAsia2016/Abstracts/Blower_abstract.pdf}{Random Matrices EurAsia 2016}, University of Macau, Macau, July 2016.
\item 
  \href{http://home.kias.re.kr/psec/nspcs16/}{The $7^{\text{th}}$ KIAS Conference on Statistical Physics}, KIAS, Seoul, Korea, July 2016.
\item
  \href{http://prcm.math.snu.ac.kr/}{The $7^{\text{th}}$ Pacific Rim Conference on Mathematics}, Seoul National University, Seoul, Korea, June 2016.
\item
  \href{http://ims-aprm2016.sta.cuhk.edu.hk/}{The $4^{\text{th}}$ Institute of Mathematical Statistics Asia Pacific Rim Meeting (IMS-APRM)}, The Chinese University of Hong Kong, Hong Kong, China, June 2016.
\item
  \href{http://icnwta4.csp.escience.cn/dct/page/1}{The $4^{\text{th}}$ International Conference on Nonlinear Waves: Theory and Applications}, Tsinghua University, Beijing, China, June 2016.
\item
  \href{http://ymsc.tsinghua.edu.cn/sanya/2016/MHTY2016/synopsis_and_organizers.aspx}{Workshop associated to Master Lectures on the Current Topics in Mathematical Physics and Probability: Horng-Tzer Yau}, Sanya, China, December 2015.
\item
  \href{http://www.birs.ca/events/2015/5-day-workshops/15w5022}{Orthogonal and Multiple Orthogonal Polynomials}, Oaxaca, Mexico, August 2015.
\item
  \href{http://www.fst.umac.mo/conference/isaac2015/}{$10^{\text{th}}$ International ISAAC Congress}, University of Macau, Macau, China, August 2015.
\item
  \href{http://www.math.pku.edu.cn/teachers/liuyong/2015forum/admin/html/default.php}{2015 Peking University Youth Probability Forum}, Peking University, Beijing, China, July 2015.
\item
  \href{http://www.2015imschina.com/}{2015 IMS-China International Conference on Statistics and Probability}, Kunming, China, July 2015.
\item
  Probability Day, Institute for Mathematical Sciences, Singapore, September 2014.
\item \href{http://th-www.if.uj.edu.pl/~matrix2014/index.html}{$30^{\text{th}}$ International Colloquium on Group Theoretical Methods in Physics}, Ghent, Belgium, July 2014.
\item
  \href{http://th-www.if.uj.edu.pl/~matrix2014/index.html}{Random Matrix Theory: Foundations and Applications}, Krakow, Poland, July 2014.
\item
  Seminar in Department of Mathematics, University of Macau, Macau, China, May 2014.
\item
  \href{http://www.maths.usyd.edu.au/u/olver/conferences/RMT.html}{Sydney Random Matrix Theory Workshop}, University of Sydney, Sydney, Australia, January 2014. 
\item
  \href{http://www.math.pku.edu.cn/teachers/liuyong/seminarC.htm}{Probability Seminar}, Peking University, Beijing, China, September 2013.
\item
  \href{http://web.eecs.umich.edu/~rajnrao/rmt2013/}{Workshop on Random Matrices and Applications}, University of Michigan, Ann Arbor, USA, June 2013.
\item
  \href{http://www6.cityu.edu.hk/ma/maicata/}{International Conferences on Approximation Theory and Applications}, City University, Hong Kong, China, May 2013.
\item
  \href{http://ww1.math.nus.edu.sg/seminars.aspx?CatID=9}{Probability Seminar}, National University of Singapore, Singapore, April 2013.
\item
  Working Group in Integrable Systems, Tsinghua University, Beijing, China, December 2012.
\item
  \href{http://www2.ims.nus.edu.sg/Programs/012random/wk2.php}{Random Matrix Theory, Applications in Statistics}, Institute for Mathematical Sciences, Singapore, Auguest 2012.
\item
  Working Group in Integrable Systems, Tsinghua University, Beijing, China, July 2012.
\item
  \href{http://mathdata.msu.edu/Seminar/Recent.asp?s=928}{Probability Seminar}, Michigan State University, Lansing, USA, April 2012.
\item
  \href{http://www.math.lsa.umich.edu/seminars_events/events.php?eventdefid=43&dt_begin=2011-07-01&dt_end=2011-12-31}{Working Group in Integrable Systems and Random Matrix Theory}, University of Michigan, Ann Arbor, USA, September 2011.
\item
  \href{http://web.hku.hk/~jeffyao/ss/}{Random matrix theory and high-dimensional statistics},  French-Chinese summer school, Changchun, China, July 2011.
\item
  \href{http://www.esf.org/activities/esf-conferences/details/2011/confdetail369.html}{Completely Integrable Systems and Applications}, Erwin Schr\"{o}dinger Institute, Vienna, Austria, July 2011.
\item
  \href{http://wis.kuleuven.be/analyse/seminar-classical.html}{Semiclassical Analysis Seminar}, Katholieke Universiteit Leuven, Leuven, Belgium, March 2011.
\item
  \href{http://www.math.lsa.umich.edu/seminars/scv/}{Several Complex Variable Seminar}, University of Michigan, Ann Arbor, USA, November 2010.
\item
  \href{http://www.math.utoronto.ca/cms/toronto-probability-seminar/}{Toronto Probability Seminar}, Fields Institute, Toronto, Canada, July 2010.
\item
  \href{http://www2.warwick.ac.uk/fac/sci/statistics/crism/workshops/orthogonal-polynomials/}{Workshop on Orthogonal Polynomials, Applications in Statistics and Stochastic Processes}, Warwick University, Coventry, UK, July 2010.
\item
  \href{http://www.math.harvard.edu/cgi-bin/showtalk.pl}{Random Matrix and Probability Seminars}, Harvard University, Cambridge, USA, April 2010.
\item
  Working Seminar in Random Matrix Theory, Peking University, Beijing, China, July 2009.
\item
  \href{http://www.crm.umontreal.ca/cal/en/mois200804.html}{Working Seminar Mathematical Physics}, Centre de Recherches Math\'{e}matiques, Montreal, Canada, April 2009.
\item
  \href{http://www.crm.umontreal.ca/cal/en/mois200811.html}{Working Seminar Matrices al\'{e}atoires, processus al\'{e}atoires, syst\`{e}mes int\'{e}grables}, Centre de Recherches Math\'{e}matiques, Montreal, Canada, November 2008.
\item
  \href{http://www.birs.ca/events/2008/5-day-workshops/08w5017}{Conference on Random Matrices, Inverse Spectral Methods and Asymptotics}, Banff International Research Station, Banff, Canada, October 2008.
\item
  \href{http://www.crm.umontreal.ca/cal/en/mois200810.html}{Seminar Physique Math\'{e}matique}, Centre de Recherches Math\'{e}matiques, Montreal, Canada, October 2008.
\item
  \href{http://www.math.lsa.umich.edu/~millerpd/WGISA_archives/WGISA_Fall_07.html}{Working group in Integrable Systems and Asymptotics}, University of Michigan, Ann Arbor, USA December 2007.
\item
  \href{http://magelhaes.hzs.be/RIMMP/}{Conference on Random and Integrable Models in Mathematics and Physics} (poster), Belgian Academy of Science, Brussels, Belgium, September 2007.
\end{item_list}

\section*{Professional Activities}

\begin{item_list}
\item Reviewer for \textit{Mathematical Reviews}, referee for \textit{Acta Math.}, \textit{Ann.\ Appl.\ Probab.}, \textit{Ann.\ Inst.\ Henri Poincar\'{e}}, \textit{Ann.\ Probab.}, \textit{Constr. Approx.}, \textit{Int.\ Math.\ Res.\ Not.\ IMRN}, \textit{J.\ Stat.\ Phys.}, \textit{J.\ Multivariate Anal.}, \textit{Peking Math.\ J.}, \textit{Phys.\ D}, \textit{Punjab Univ.\ J.\ Math.}, \textit{Random Matrices Theory Appl.}, \textit{SIAM J.\ Matrix Anal.\ Appl.}, \textit{SIGMA}, and \textit{Statist.\ Probab.\ Lett.}.
\item Co-organizer of the workshop:
  \href{https://ims.nus.edu.sg/events/2020/random/index.php}{Random Matrix EurAsia 2020}, Institute for Mathematical Sciences, Singapore, May 2020.
\item Co-organizer of the workshop:
  \href{http://www2.ims.nus.edu.sg/Programs/015wrandom/index.php}{Workshop on Stochastic Processes in Random Media}, Institute for Mathematical Sciences, Singapore, May 2015.
\item Co-organizer of the seminar:
  \href{http://www.math.lsa.umich.edu/seminars_events/events.php?eventdefid=43&dt_begin=2011-07-01&dt_end=2011-12-31}
  {Working Group in Integrable Systems and Random Matrix Theory}, University of Michigan, 2011--2012.
\end{item_list}

\section*{Grants}

\begin{item_list}
\item 
  Chinese NSFC grant 11871425.
\item
  Singapore AcRF Tier 1 Grant, R-146-000-262-114.
\item
  Singapore AcRF Tier 1 Grant, R-146-000-217-112.
\item
  Start-up grant in National University of Singapore, R-146-000-164-133.
\end{item_list}

% \section*{Conferences and Workshops Attended}

% \begin{item_list}
% \item \href{http://www.msri.org/web/msri/scientific/workshops/show/-/event/Wm510}{Random Matrix Theory and its Applications II},
%   Mathematical Sciences Research Institute, December 2010.
% \item \href{http://www.msri.org/web/msri/scientific/workshops/show/-/event/Wm501}{Random Matrix Theory and its Applications I},
%   Mathematical Sciences Research Institute, September 2010.
% \item \href{http://www.crm.umontreal.ca/Tilings08/index_e.shtml}{Random Tilings, Random Partitions and Stochastic Growth Processes}, Centre de Recherches Math\'{e}matiques, September 2008.
% \item \href{http://www.crm.umontreal.ca/Matrices08/index_e.shtml}{Random Matrices, Related Topics and Applications}, Centre de Recherches Math\'{e}matiques, August 2008.
% \item \href{http://www.crm.umontreal.ca/Laplacian08/index_e.shtml}{Laplacian Growth and Related Topics}, Centre de Recherches Math\'{e}matiques, August 2008.
% \item \href{http://www.crm.umontreal.ca/Loewner08/schedule.shtml}{Stochastic Loewner Evolution and Scaling Limits}, Centre de Recherches Math\'{e}matiques, August 2008.
% \item Joint Summer Research Conference on Interactions of Random Matrix Theory, Integrable Systems, and Stochastic Processes, Snowbird, UT, June 2007.
% \item \href{http://www.aimath.org/conferences/ntrmt/}{Number Theory and Random Matrix Theory}, University of Rochester, June 2006.
% \item \href{http://math.arizona.edu/~mcl/ISRMA.html}{Integrable Systems, Random Matrices, and Applications}, Courant Institute, May 2006.
% \end{item_list}

%\section*{Miscellaneous}

%\begin{item_list}
%\item \textit{Security Clearance:} Special Sworn Status, U.S. Census Bureau.
%\end{item_list}

% Footer
\begin{center}
  \begin{small}
    Last updated: \today
  \end{small}
\end{center}

\end{document}

\section*{References}

\begin{minipage}[t]{0.5\linewidth}
  Jinho Baik \\
  Department of Mathematics \\
  University of Michigan \\
  Ann Arbor, MI 48109 \\
  (734) 763-2150 \\
  \url{baik@umich.edu}
\end{minipage}
\begin{minipage}[t]{0.5\linewidth}
  Mark Adler \\
  Department of Mathematics \\
  Brandeis University \\
  Waltham, MA 02453 \\
  (781) 736-3055 \\
  \url{adler@brandeis.edu}
\end{minipage}

\bigskip

\begin{minipage}[t]{0.5\linewidth}
  Pierre van Moerbeke \\
  D\'{epartment} de Math\'{e}matiques \\
  Universit\'{e} catholique de Louvain \\
  1348 Louvain-la-Neuve, Belgium \\
  +32 (0)10 47 31 52 \\
  \url{pierre.vanmoerbeke@uclouvain.be}
\end{minipage}
\begin{minipage}[t]{0.5\linewidth}
  Craig Tracy \\
  Department of Mathematics \\
  University of California \\
  Davis, CA 95616 \\
  (530) 564-0575 \\
  \url{tracy@math.ucdavis.edu}
\end{minipage}

\bigskip

\begin{minipage}[t]{0.5\linewidth}
  Karen Rhea (teaching) \\
  Department of Mathematics \\
  University of Michigan \\
  Ann Arbor, MI 48109 \\
  (734) 936-9936 \\
  \url{krhea@umich.edu}
\end{minipage}

\medskip

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
