\documentclass[a4paper]{article}

\usepackage{url, fullpage, hyperref}
\usepackage{enumitem}

\title{Syllabus of Measure and Integration}
\date{}

\begin{document}
\maketitle

\begin{minipage}[t]{0.5\linewidth}
  \begin{itemize}[leftmargin=*]
  \item Lecture time: Monday, Thursday 12:00--14:00
  \item Lecture room: LT34
  \item Tutorial time: Wednesday 12:00--13:00, 13:00--14:00
  \item Tutorial room: Block S17, Room 04--05 
  \end{itemize}
\end{minipage}
\begin{minipage}[t]{0.4\linewidth}
  \begin{itemize}
  \item Lecturer and tutor: A/Prof Wang Dong
  \item Office: Block S17, Room 06--20
  \item Tel: 6516\ 2746
  \item Email: \href{mailto:matwd@nus.edu.sg}{\nolinkurl{matwd@nus.edu.sg}}
  \end{itemize}
\end{minipage}

\begin{itemize}
\item Prerequsite: MA3209
  
\item Course description: This module is suitable not only for mathematics majors, but also for science and engineering majors who need a rigorous introduction to the concepts of measures and integrals. It covers Lebesgue measure and Lebesgue integral in a rigorous manner. We begin complicated proofs with an introduction which shows why the proof works. Examples are included to show why each hypothesis of a major theorem is necessary. Major topics: Lebesgue measure. Outer measure. Measurable sets. Regularity of Lebesgue measure. Existence of non-measurable sets. Measurable functions. Egoroff’s Theorem. Lusin’s Theorem. Lebesgue integral. Convergence theorem. Differentiation. Vitali covering lemma. Functions of bounded variation. Absolute continuity. $L^p$ spaces. Holder’s inequality. Minkowski’s inequality. Riesz--Fischer theorem.
  
% \item Course centents:
%   \begin{itemize}
%   \item 
%     Basics of normed vector spaces
%   \item 
%     Linear operators in normed vector spaces
%   \item 
%     Basics of Hilbert spaces
%   \item 
%     Linear operators in Hilbert spaces
%   \end{itemize}
\item References:
  \begin{itemize}
  \item
    R.G.~Bartle, \emph{The elements of integration and Lebesgue measure}, Wiley, 1995. (Main textbook)
  \item    
    H.L.~Royden and P.M.~Fitzpatrick, \emph{Real analysis} (4th edition), Prentice Hall, 2010.
  \item
    D.W.~Strook, \emph{A concise introduction to the theory of integration}, World Scientific, 1990.
  \item
    E.M.~Stain and R.~Shakarchi, \emph{Real analysis: measure theory, integration, and Hilbert spaces}, Princeton, 2005.
  \end{itemize}
\item Assessment:
  Assessment of students will be based on
  \begin{itemize}
  \item a one-hour test during lecture time (\emph{tentatively} on 1 October 2020), 25\%
  \item tutorial participation, 5\%,
  \item four sets of homework, 10\%,
  \item a two-and-a-half hour final examination, 60\%.
  \end{itemize}
  \emph{Any student who is absent without a valid reason from an assessment will be given zero mark for that assessment.}
\end{itemize}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
